---
title: "Functions to handle errors in R and python"
author: "Pep Porrà"
date: '2022-11-26'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goal

Explore functions to handle errors (intercept them) in both R and python

## References

* Advance R
* Efficient R programming
* Python for dats Scientists
* Python handbook

## R

Let us take a simple example. We call a function `f()` that, in some cases, 
raises an error and stops execution. If `f()` function
is called within one of our functions or scripts, it will stop as soon as `f()`
fails. 

```{r}
f <- function(x) charToRaw(x)
```


```{r, eval = FALSE}
f(9)
#> Error in charToRaw(x) : argument must be a character vector of length 1
```

```{r, eval = FALSE}
our_function <- function(x){
  y <- f(x)
  now <- Sys.time()
  class(y)
  return(now)
}

our_function(9)
#> Error in charToRaw(x) : argument must be a character vector of length 1
```

### `try()`


```{r}
our_function_1 <- function(x){
  y <- try(f(x))
  now <- Sys.time()
  print(class(y))
  return(now)
}
our_function_1(9)
```

With `try()` function, execution does not stop after the error. Error message is
shown but code after the error is executed


```{r}
our_function_1 <- function(x){
  y <- try(f(x), silent = TRUE)
  now <- Sys.time()
  class(y)
  return(now)
}
our_function_1(9)
```
Parameter `silent = TRUE` of `try()` function prevents the message error to be
shown.

Note that `try()` return an object of class "try-error" when `f()`call fails,
which can be used to run specific actions after the error [1]:

```{r}
our_function_1 <- function(x){
  y <- try(f(x), silent = TRUE)
  if (class(y) == "try-error") {
    # action after the error is detected
    message("ERROR in function f")
  }
  now <- Sys.time()
  class(y)
  return(now)
}
our_function_1(9)
```

### `tryCatch()`

```{r}
our_function_2 <- function(x){
  y <- tryCatch( f(x),
    error = function(c) {
      message("standard error message: ", c$message, "\n")
      message("Additional information ...\n")
    },
    finally = message("End of function tryCatch()"))
  if (class(y) == "try-error") {
    # action after the error is detected
    message("ERROR in function f")
  } else {
    message(paste("y =", y))
  }
  now <- Sys.time()
  message(paste("class of y:", class(y)))
  return(now)
}
our_function_2(9)
```
```{r}
suppressMessages(our_function_2(9))
```

```{r}
our_function_2("9")
```

## Python

```{r}
reticulate::py_config()
```

```{python}
# def 
```

